Approximation as Differentiation

### 2025-07-08 main/models.tex
Minor corrections.

\begin{example}
  The following categories are $\CMon$-enriched and have finite
  products, and hence biproducts:
  \begin{enumerate}
  \item In $\FinVect$, morphisms are linear maps and so can be added
    and have a zero map. Finite products are given by cartesian
    products of the underlying sets, with the vector operations
    defined pointwise.
  \item In $\LatGal$, right adjoints are summed using meets and left
    adjoints are summed using joins. The zero maps are given by the
    constantly $\top$ and constantly $\bot$ functions
    respectively. Products are given by the cartesian product of the
    underlying set.
  \item The $\LatGal$ example splits into two separate categories
    $\MeetSLat$ and $\JoinSLat$ of meet semilattices and join
    semilattices respectively, with their respective homomorphisms as
    the morphisms. As for $\LatGal$, these categories are
    $\CMon$-enriched and have finite products, and so finite
    biproducts.
  \end{enumerate}
\end{example}

@@

\begin{proposition}
  The functor
  $F : \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$ defined
  on objects as
  $F(X, \partial X) = (X, \lambda x. (\partial X(x), \partial
  X(x)))$. This functor is faithful and preserves coproducts and
  finite products.
\end{proposition}

With this embedding functor, we can see that the interpretation of
first-order types will be the same up to isormorphism in
$\Fam(\LatGal)$ and $\Fam(\MeetSLat \times \JoinSLat^\op)$, as long as
we interpret the base types as objects in $\Fam(\LatGal)$. At
higher-order, however, the meet-semilattice and join-semilattice
sides of the interpretation will diverge, and it is no longer clear
that the interpretation of programs using higher-order functions
internally will result in Galois connections. In \secref{definability}
we will see that every program with first-order type (even if it uses
higher-order functions internally) does in fact have an interpretation
definable in $\Fam(\LatGal)$.

>>>>

\begin{example}
  The following categories are $\CMon$-enriched and have finite
  products, and hence biproducts:
  \begin{enumerate}
  \item In $\FinVect$, morphisms are linear maps and so can be added
    and have a zero map. Finite products are given by cartesian
    products of the underlying sets, with the vector operations
    defined pointwise.
  \item In $\LatGal$, right adjoints are summed using meets and left
    adjoints are summed using joins. The zero maps are given by the
    constantly $\top$ and constantly $\bot$ functions
    respectively. Products are given by the cartesian product of the
    underlying set.
  \item The $\LatGal$ example splits into two separate categories
    $\MeetSLat$ and $\JoinSLat$ of meet semilattices and join
    semilattices respectively, with their respective homomorphisms as
    the morphisms. As with $\LatGal$, these categories are
    $\CMon$-enriched and have finite products, and so finite
    biproducts.
  \end{enumerate}
\end{example}

@@

\begin{proposition}
  The functor
  $F : \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$ is defined
  on objects as
  $F(X, \partial X) = (X, \lambda x. (\partial X(x), \partial
  X(x)))$. This functor is faithful and preserves coproducts and
  finite products.
\end{proposition}

With this embedding functor, we can see that the interpretation of
first-order types will be the same up to isomorphism in
$\Fam(\LatGal)$ and $\Fam(\MeetSLat \times \JoinSLat^\op)$, as long as
we interpret the base types as objects in $\Fam(\LatGal)$. At
higher-order, however, the meet-semilattice and join-semilattice
sides of the interpretation will diverge, and it is no longer clear
that the interpretation of programs using higher-order functions
internally will result in Galois connections. In \secref{definability}
we will see that every program with first-order type (even if it uses
higher-order functions internally) does in fact have an interpretation
definable in $\Fam(\LatGal)$.
